Peer Assessment 1 - Reproducible Research Course
========================================================

Loading and preprocessing the data
--------------------------------------
Load the data setting the correct directory where the file is.  
Read the .csv file
```{r echo=TRUE}
setwd("~/Curso 5 Reproducible research/CP1")
data<-read.csv("activity.csv")
```

Process the data in order to remove the NA data
```{r echo=TRUE}
good<-complete.cases(data)
data1<-data[good,]
data1$date<-factor(data1$date)
```

What is mean total number of steps taken per day?
-----------------------------------------------------

histogram of the total number of steps taken each day.
```{r echo=TRUE}
stepsday<-tapply(data1$steps, data1$date, sum)
hist(stepsday, main="Nº of steps")
```

Mean and median total number of steps taken per day
```{r echo=TRUE}
means<-tapply(data1$steps,data1$date,mean)
medians<-tapply(data1$steps,data1$date,median)
```

* mean steps per day
```{r echo=TRUE}
means
```
* median steps per day
```{r echo=TRUE}
medians
```

What is the average daily activity pattern?
---------------------------------------------------
```{r echo=TRUE}
data1$interval<-as.factor(data1$interval)
averagesteps<-tapply(data1$steps, data1$interval, mean)
plot(c(1:288),averagesteps, type="l", xlab="intervals", main="Average daily activity pattern", col="blue")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r echo=TRUE}
y<-cbind(factor(data1$intervals), averagesteps)
which.max(y[,1])
```
* The maximum number of steps is on interval 835

Imputing missing values
------------------------------

New dataset that is equal to the original dataset but with the missing data filled in.     
* Strategy used to fill missing data: replace NA values with the mean of the 5-minute interval missed  

Replace the NA values with the average daily activity pattern obtained before to create a new dataset:
```{r echo=TRUE}
data2<-replace(data, is.na(data), averagesteps)
```

Histogram of the total number of steps taken each day
```{r echo=TRUE}
stepsday2<-tapply(data2$steps, data2$date, sum)
hist(stepsday2, main="Nº of steps")
```

Mean and median total number of steps taken per day
```{r echo=TRUE}
means2<-tapply(data2$steps,data2$date,mean)
medians2<-tapply(data2$steps,data2$date,median)
```

* mean steps per day
```{r echo=TRUE}
means2
```
* median steps per day
```{r echo=TRUE}
medians2
```


Are there differences in activity patterns between weekdays and weekends?
-------------------------------------------------------------------------------
Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r echo=TRUE}
Sys.setlocale("LC_TIME", "en")
weekdays<-ifelse((weekdays(as.Date(data2$date))) %in% c("Saturday", "Sunday"), "weekend","weekday")
data2$weekday<-as.factor(weekdays)
```

Panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
```{r echo=TRUE}
averagesteps2<-data.frame(tapply(data2$steps, list(data2$interval,data2$weekday), mean))
par(mfrow=c(2,1))
plot(row.names(averagesteps2),averagesteps2$weekend, type="l", ylab= "Number of steps", xlab="Interval", col="blue", main="Weekends")
plot(row.names(averagesteps2),averagesteps2$weekday, type="l", ylab= "Number of steps", xlab="Interval", col="blue", main="Weekdays")
```

